---
title: "Analysis of the UCH dataset with LMWiRe"
author: "Franceschini Sébastien"
date: "`r Sys.Date()`"
package: PepsNMR
output:
 BiocStyle::html_document:
    toc: yes
    toc_depth: 3
    toc_float: yes
vignette: >
  %\VignetteIndexEntry{Analysis of the UCH dataset with LMWiRe}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::knitr}
editor_options: 
  chunk_output_type: console
---

# Introduction

The main aim of this vignette is to give a brief example of the LMWiRe package. This package has been created to analyze models with wide response and categorical parameters. This example present all the usual step of the analysis starting from the data importation to the result visualisation. 

# Installation

The package is actually in its developping stage and is available on GitHub at the url : <https://github.com/FranceschiniS/LMWiRe>.It can be installed via the `remotes::install_github()` function. The pacakage needs to be loaded afterwards.


```{r Install, results=FALSE, message=FALSE}

remotes::install_github("FranceschiniS/LMWiRe")

library("LMWiRe")

```

# Data importation

Before any analysis the `UCH data` needs to be loaded. The LMWiRe package contains the dataset and it can be load with the `data()` function. The `load()` function is also useful to import your own data.

```{r Data Importation}

load("UCH.RData")

```

# Data visualisation

The UCH dataset stands for Urine-Citrate-Hippurate and comes from \textit{Rousseau et al.} It contains 3 elements : a matrix `outcomes` with 34 observations of 600 response variables representing the spectra from the H-NMR spectroscopy, a formula for the GLM model used and `design` matrix with 34 observations and 5 explanatory variables. It is a good habit to start by a data visualisation.

First the 600 response variables are the spectrum of the H-NMR for each observations. They can be plot together with the simple `geom_line` function from the `ggplot2` package.

```{r Spectrum visualization}
firstobslab = names(UCH$outcomes[,1])[1]

ggplot2::ggplot(data = as.data.frame(t(UCH$outcomes))) + 
  ggplot2::geom_line(ggplot2::aes(x = as.numeric(colnames(UCH$outcomes)), y = UCH$outcomes[1, ])) +
  ggplot2::scale_x_reverse(lim = c(10, 0)) +
  ggplot2::xlab("ppm") +
  ggplot2::ylab("Intensity")+
  ggplot2::ggtitle(paste("H-NMR spectrum for the first observation : ",firstobslab))+
  ggplot2::theme(plot.title = ggplot2::element_text(hjust = 0.5))

```

Secondly the experimental design can be plotted. The function `PlotDesign` plot a graph of the experimental design for 3 differents variables. For more variables a scatterplot is necessary.

```{r Design Visualization, warning=FALSE}

PlotDesign(design = UCH$design, var1_name = "Citrate",var2_name = "Hippurate",var3_name = "Day")

```

# PCA (utile?)

The function `SVDforPCA` is useful to compute a PCA decomposition of the `outcomes` matrix. The `scores` and `loadings` can then be plotted with the function `DrawScores` and `DrawLoadings`.

```{r PCA}
ResPCA = SVDforPCA(UCH$outcomes)
```

```{r VariancesPercentages}

eig.res = rbind(ResPCA$var[1:6], ResPCA$cumvar[1:6])
rownames(eig.res) = c("Variances", "Cum Var Values")
pander::pander(eig.res)
pander::pander("A scree plot should be perform")

```


```{r Scores}

DrawScores(ResPCA, type.obj = "PCA", drawNames = TRUE, createWindow = F, 
    main = "Reponse matrix score plot", color = UCH$design$Citrate, pch = UCH$design$Hippurate, 
    axes = c(1, 2), size = 2.5) + ggplot2::scale_color_discrete(name = "Citrate") + ggplot2::scale_shape_discrete(name = "Hippurate")

```

```{r Loadings}
DrawLoadings(ResPCA,type.obj = "PCA", main="Loadings from the two first component")
```

# GLM decomposition

```{r ModelMatrix}

ResLMModelMatrix = LMModelMatrix(UCH$formula,UCH$design)

```

```{r EffectMatrices}

ResLMEffectMatrices = LMEffectMatrices(ResLMModelMatrix,UCH$outcomes)

```

```{r}
ResLMSS = LMSS(ResLMEffectMatrices)
```

```{r}
PlotLMSS(ResLMSS = ResLMSS,abbrev=TRUE)
```


# ASCA

```{r}
# ResASCA = LMWiRe::PCALMEffects(ResLMEffectMatrices = ResLMEffectMatrices,method="ASCA")
```

```{r}
# DrawScores(ResASCA,type.obj = "PCA")
```

# ASCA-E

# APCA

# Session info
